---
title: "Functional Visualization Approach"
output: html_document
---

# Retrieve the data

Link to file:

```{r}
wideLink <- "https://github.com/MagallanesAtAlacip/datafiles/raw/main/covid_fatalities2021_yearly_wide.csv"
```

Reading from link:

```{r}
covidPeru_wide <- read.csv(wideLink)
head(covidPeru_wide)
```

```{r}
str(covidPeru_wide)
```

As you see, each row is a province. Then each column as a specific count. for example:

> The first cell next to ABANCAY tells there were **15** fatalities which were less than 40 in the year 2020.

A functional approach can work well with this:

```{r}
# 2. Prepare numeric wide data (Drop the first column "PROVINCIA")
numeric_data <- covidPeru_wide[, -1]

boxplot(numeric_data,
        names = names(numeric_data))
```

The `boxplot()` function accepts more arguments;

```{r}
boxplot(numeric_data,
        horizontal = TRUE,
        las = 1)
```

Logging 'x':

```{r}

boxplot(numeric_data,
        horizontal = TRUE,                            # horizontal=TRUE is vert=False
        log = "x",                                    # Appies log scale to X-axis
        las = 1,                                      # las=1 keeps text horizontal (readable)
        xlab = "Count")
```

R base plots do not handle log(0) well; We need the `+1` trick:

```{r}
# Add 1 to all your data to eliminate the 0 problem safely
logged_data <- numeric_data + 1

# Plot using the modified data
boxplot(logged_data,
        horizontal = TRUE,
        las = 1,
        log = "x",      # R can now scale it perfectly
        col = "skyblue")     # Hide the axis so we can label it nicely
```



