---
title: "Social Network Viz"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = "center")
rm(list = ls())
```

We have a real Twitter interaction network among **46** people in the Seattle tech/startup scene: **393** undirected ties, weighted by `traffic` (interaction count), with two node attributes — `male` (binary) and `followers` (follower count).


# Open the files

We generally need two (edges and node attributes).

```{r load-libs}
library(igraph)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggraph)
```

```{r load-data}
fileLink_edges      <- "https://github.com/MagallanesAtAlacip/datafiles/raw/main/edges.csv"
fileLink_attributes <- "https://github.com/MagallanesAtAlacip/datafiles/raw/main/nodeattributes.csv"

edges      <- read.csv(fileLink_edges)
attributes <- read.csv(fileLink_attributes)
```

```{r show-edges}
edges|>head()
```

```{r show-attrs}
attributes |>head()
```


# Creating the Network from DataFrames

Use the previous data frames to create the network. `graph_from_data_frame()` takes the edge list as the
first argument and the node attribute table as `vertices` — this attaches `male` and `followers` to every
node in one step, instead of the two-step dictionary approach some tools require.

```{r build-graph}
EliteNet <- graph_from_data_frame(edges, directed = FALSE, vertices = attributes)
EliteNet
```

**A scoping decision:** `traffic` is stored as an edge attribute, but density, diameter, centrality, and
community detection below are computed on the **unweighted** graph. This talk is about topological
structure — who is reachable from whom, who bridges gaps, whether communities exist — and mixing in
interaction volume would answer a different question ("who talks most"). Weighting by `traffic` is a
legitimate follow-up analysis, just not this one.

A view of edges and node attributes:

```{r view-edges}
head(igraph::as_data_frame(EliteNet, what = "edges"), 5)
```

```{r view-nodes}
head(igraph::as_data_frame(EliteNet, what = "vertices"), 5)
```

A simple call to `plot()` will draw the graph:

```{r naive-draw}
plot(EliteNet, vertex.label = NA, vertex.size = 5, edge.color = "grey80")
```

This is the naive first look — no color, no attributes, no seed. It confirms a graph exists and nothing
else. Every plot from here on earns its place by testing a specific claim.

# Exploration of the network

## Basic metrics

### Connectedness

A network is "connected" if there exists a *path* between any pair of nodes (undirected networks).

```{r connected}
is_connected(EliteNet)
```

> **TRUE** — all 46 people are mutually `reachable`. That matters mechanically, too: `diameter()` on a
disconnected graph in igraph returns `Inf` for the unreachable pairs rather than erroring, which would
silently give a misleading "worst case" number — so this check is what makes the diameter figure below
trustworthy, not just a formality.

### Density

From 0 to 1, where 1 makes it a "complete" network: there is a link between every pair of nodes.

```{r density}
edge_density(EliteNet)
```

> **0.38.** Of the 1,035 possible pairs among 46 people, 393 (38%) are actually connected. For a social
network at this scale, that's dense — this reads less like a loose follower graph and more like a tight,
high-contact group who mostly already know each other.

### Diameter

When two vertices are connected, one can reach the other using multiple edges. The geodesic is the
shortest path between two connected vertices. The diameter is the maximum geodesic in a network.

```{r diameter}
diameter(EliteNet)
```

> That is the worst-case scenario for a person to reach another: having `diameter - 1` intermediaries.

## Visuals

Visuals are key for complementary analysis. But it is not an easy task.

### Assortativity

Assortativity measures whether nodes tend to connect to other nodes similar to themselves. Closer to 1
means higher assortativity, closer to -1 means disassortativity, while 0 is no assortativity.
Assortativity can be measured in three complementary ways.

**a. Categorical assortativity:** tendency for nodes to connect with other nodes sharing the same
category.

```{r assort-gender}
# igraph's assortativity_nominal needs integer category codes starting at 1
gender_types <- V(EliteNet)$male + 1
assortativity_nominal(EliteNet, types = gender_types, directed = FALSE)
```

> With r ~ 0.043, gender is effectively irrelevant: same-sex and cross-sex ties are distributed no differently than chance would predict.

Layouts help show the complexity of a graph without becoming an analytical tool on their own. If patterns
are not shockingly clear, eye bias may lead us to interpret visual artifacts as if they meant something.

Let's prepare a layout to see if we can effectively interpret the close-to-zero value of assortativity
for the `male` attribute — coloring nodes by gender to see if the (lack of) pattern is visible:

```{r gender-layout}
set.seed(42)
lay <- layout_with_fr(EliteNet, niter = 2000)

plot(EliteNet, layout = lay,
     vertex.color = ifelse(V(EliteNet)$male == 1, "steelblue", "orange"),
     
     vertex.label = NA, vertex.size = 6, edge.color = "grey85")
```

A next step is to look at the edge pattern directly. Let's split the edges into same-sex and cross-sex:

```{r same-cross-sex}
el <- as_edgelist(EliteNet, names = FALSE)
male_vec <- V(EliteNet)$male

same_sex  <- el[male_vec[el[,1]] == male_vec[el[,2]], , drop = FALSE]
cross_sex <- el[male_vec[el[,1]] != male_vec[el[,2]], , drop = FALSE]

plot(EliteNet, layout = lay,
     vertex.color = ifelse(V(EliteNet)$male == 1, "steelblue", "orange"),
     vertex.label = NA, vertex.size = 6,
     edge.color = NA)

# cross-sex edges first, in grey
plot(graph_from_edgelist(cross_sex, directed = FALSE), layout = lay, add = TRUE,
     vertex.label = NA, vertex.size = 6, vertex.color = NA,
     edge.color = adjustcolor("grey60", alpha.f = 0.5))

# same-sex edges on top, in crimson
plot(graph_from_edgelist(same_sex, directed = FALSE), layout = lay, add = TRUE,
     vertex.label = NA, vertex.size = 6, vertex.color = NA,
     edge.color = adjustcolor("firebrick", alpha.f = 0.4))
```

Can you detect if gender homophily exists? 

Most of the time, visuals require knowing what math actually happened behind the scenes. In this case,
assortativity needs a simple cross-tabulation as the initial input:

```{r mixing-matrix}
male_named <- setNames(V(EliteNet)$male, V(EliteNet)$name)
el_names <- as_edgelist(EliteNet, names = TRUE)

mix <- table(male_named[el_names[,1]], male_named[el_names[,2]])
mix
```

`mix` is a raw crosstab of edge endpoints by `male` — not yet symmetrized or normalized. Because the
graph is undirected, an edge listed as (female, male) and one listed as (male, female) mean the same
thing, so we symmetrize before normalizing:

```{r mixing-matrix-sym}
mix_sym <- (mix + t(mix)) / 2
mix_sym
```

```{r mixing-matrix-norm}
M <- mix_sym / sum(mix_sym)
M
```

| Cell     | Meaning                                      |
| -------- | --------------------------------------------- |
| `M[1,1]` | Fraction of edges connecting female <-> female |
| `M[1,2]` | Fraction of edges connecting female <-> male   |
| `M[2,1]` | Fraction of edges connecting male <-> female   |
| `M[2,2]` | Fraction of edges connecting male <-> male     |

We can compute assortativity from the table:

```{r manual-assort}
a <- rowSums(M)
r_manual <- (sum(diag(M)) - sum(a^2)) / (1 - sum(a^2))
r_manual
```

| Line                | What it does                  | Mathematical meaning                                                    |
| -------------------- | ------------------------------ | ------------------------------------------------------------------------ |
| `a <- rowSums(M)`    | Row sums of the mixing matrix | $a_i = \sum_j e_{ij}$ — fraction of all edge-ends attached to group $i$   |
| `sum(diag(M))`       | Sum of diagonal elements      | $\sum_i e_{ii}$ — fraction of same-type edges                            |
| `sum(a^2)`           | Sum of squared marginals      | $\sum_i a_i^2$ — what the diagonal would be under independence           |
| `1 - sum(a^2)`       | Denominator                   | Normalization factor — maximum possible deviation from independence      |

$$r = \frac{\sum_i e_{ii} - \sum_i a_i^2}{1 - \sum_i a_i^2}$$

This matches `assortativity_nominal()` above, confirming the built-in function isn't a black box. A
simple heatmap makes the lack of assortativity visible:

```{r mixing-heatmap}
M_df <- as.data.frame(as.table(M))
colnames(M_df) <- c("row", "col", "value")

ggplot(M_df, aes(x = col, y = row, fill = value)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(value, 3)), color = "white") +
  scale_fill_viridis_c(limits = c(0, 0.5)) +
  labs(x = NULL, y = NULL, fill = "fraction") +
  theme_minimal()
```

**b. Assortativity (numerical):**

```{r assort-followers}
assortativity(EliteNet, values = V(EliteNet)$followers, directed = FALSE)
```

> The result above suggests no disassortativity by follower count. Value too close to zero.

```{r followers-layout}
followers_vec <- V(EliteNet)$followers


ggraph(EliteNet, layout = "manual", x = lay[,1], y = lay[,2]) +
  geom_edge_link(color = "grey85", alpha = 0.7) +
  geom_node_point(aes(color = followers), size = 4) +
  scale_color_continuous(palette = "viridis")+
  theme_void()
```

It is not easy to read the lack of assortativity from the plot above. Again: reflect on what we're
representing mathematically. The question is `do popular people tend to connect with popular people?`
or `with unpopular ones?` Let's look at the edge behavior directly:

```{r followers-pairs}
pairs_df <- data.frame(
  u_fol = followers_vec[el[,1]],
  v_fol = followers_vec[el[,2]]
)

cor(pairs_df$u_fol, pairs_df$v_fol)
```

Note: unlike a directed edge list from a pandas-style source, `igraph::assortativity()` already treats
the graph as undirected internally and symmetrizes both edge directions before correlating — so the
number above from a naive `u`/`v` correlation on a single, arbitrarily-ordered edge list can differ
slightly from `assortativity()`'s own result. To see the "true" symmetric version by hand, add both
directions explicitly, exactly like the built-in function does:

```{r followers-pairs-sym}
pairs_sym <- rbind(
  data.frame(u_fol = followers_vec[el[,1]], v_fol = followers_vec[el[,2]]),
  data.frame(u_fol = followers_vec[el[,2]], v_fol = followers_vec[el[,1]])
)

cor(pairs_sym$u_fol, pairs_sym$v_fol)
```

Now we can use `pairs_sym` for the scatter plot:

```{r followers-scatter}
ggplot(pairs_sym, aes(x = u_fol, y = v_fol)) +
  geom_point(alpha = 0.25, color = "grey40") +
  geom_smooth(method = "lm", color = "darkorange", se = TRUE) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", alpha = 0.5) +
  labs(x = "u followers", y = "v followers",
       title = "Follower count of connected pairs") +
  theme_minimal()
```

**c. Degree assortativity:** connections based on a different kind of popularity — becoming a hub or a
spoke.

```{r assort-degree}
assortativity_degree(EliteNet, directed = FALSE)
```

> The result above suggests disassortativity: poorly-connected nodes tend to connect with
> highly-connected ones.

```{r degree-layout}
deg <- degree(EliteNet)

ggraph(EliteNet, layout = "manual", x = lay[,1], y = lay[,2]) +
  geom_edge_link(color = "grey85", alpha = 0.7) +
  geom_node_point(aes(color = deg, size = deg), alpha = 0.7) +
  scale_color_continuous(palette = "viridis") +
  scale_size(range = c(3, 12),guide = NULL) + #not this legens
  theme_void()



```

We can use degree the same way we used followers as a node attribute:

```{r degree-excess}
deg_u <- deg[el[,1]] - 1
deg_v <- deg[el[,2]] - 1

pairs_excess <- rbind(
  data.frame(u_excess = deg_u, v_excess = deg_v),
  data.frame(u_excess = deg_v, v_excess = deg_u)
)

cor(pairs_excess$u_excess, pairs_excess$v_excess)
```

When computing degree assortativity, we subtract 1 because the edge you're standing on is already spent.
The question is not "do popular people have popular friends?" but "given this friendship, do their
*remaining* friendships cluster?" The `-1` accounts for the connection you've already used to get there.


```{r}
library(ggExtra)

p <- ggplot(pairs_excess, aes(x = u_excess, y = v_excess)) +
  geom_point(alpha = 0.3, size = 1.5, color = "steelblue") +
  geom_smooth(method = "lm", color = "firebrick") +
  theme_minimal()

ggMarginal(p, type = "histogram", fill = "steelblue", color = "white")
```


```{r degree-excess-plot}
ggplot(pairs_excess, aes(x = u_excess, y = v_excess)) +
  geom_hex(bins = 15) +
  scale_fill_viridis_c(option = "viridis", direction = -1) +
  labs(title = "Excess degree of connected pairs") +
  theme_minimal()
```

# Exploration of network nodes

- **Eigenvector centrality** tells you how well-connected a vertex is; vertices with the highest values
  are the most influential, as they connect to other well-connected vertices.
- **Closeness** tells you how close a vertex is to every other vertex. A vertex with high closeness can
  share information faster than the rest.
- **Betweenness** tells you how critical a vertex is for connecting vertices that aren't directly
  connected.

```{r centrality}
eig  <- eigen_centrality(EliteNet)$vector
clos <- closeness(EliteNet)
betw <- betweenness(EliteNet)

DFCentrality <- data.frame(
  person      = V(EliteNet)$name,
  Eigenvector = eig,
  Closeness   = clos,
  Betweenness = betw
)

head(DFCentrality)
```

One note on the plot below: raw eigenvector values span a wide range and, plotted linearly, small
differences near the bottom disappear. Rather than an unexplained exponent, we min-max scale eigenvector
centrality to a fixed marker-size range (2-10) — the transform is linear and documented, so size
differences you see are proportional to the underlying centrality, not an artifact of the scaling choice.

```{r centrality-scatter}
eig_min <- min(DFCentrality$Eigenvector)
eig_max <- max(DFCentrality$Eigenvector)
DFCentrality$size <- 2 + 8 * (DFCentrality$Eigenvector - eig_min) / (eig_max - eig_min)

ggplot(DFCentrality, aes(x = Betweenness, y = Closeness, size = size, label = person)) +
  geom_point(alpha = 0.5, color = "#4C72B0") +
  geom_text(size = 2.5, alpha = 0.6, vjust = -1) +
  scale_size_identity() +
  labs(title = "Centrality scatterplot (size = eigenvector centrality, linearly scaled)") +
  theme_minimal() +
  theme(legend.position = "none")
```

The previous results tell us that a couple of people are salient:

```{r top-hubs}
top_hubs <- DFCentrality %>% arrange(desc(Eigenvector)) %>% head(2) %>% pull(person)
top_hubs
```

Let's highlight those:

```{r hub-plot}
hub_flag <- V(EliteNet)$name %in% top_hubs

plot(EliteNet, layout = lay,
     vertex.label = ifelse(hub_flag, V(EliteNet)$name, NA),
     vertex.label.cex = 0.8, vertex.label.color = "blue",
     vertex.color = ifelse(hub_flag, "red", "grey70"),
     vertex.size = ifelse(hub_flag, 8, 4),
     edge.color = adjustcolor("grey70", alpha.f = 0.3))
```

# Communities

Before hunting for communities, we first ask: does EliteNet have more triangles than chance would
produce? If not, any "communities" we find might be algorithmic artifacts — patterns forced by the
algorithm rather than real social structure. Transitivity measures "friend of a friend is a friend" —
the raw material for communities. We compare EliteNet to a random null model with the same number of
nodes and edges. If the ratio is near 1, triangles are random noise. 

```{r transitivity-null}
elite_trans <- transitivity(EliteNet, type = "global")
cat(sprintf("EliteNet transitivity: %.4f\n", elite_trans))

n <- vcount(EliteNet)
m <- ecount(EliteNet)

set.seed(123)
random_transitivities <- replicate(100, transitivity(sample_gnm(n, m), type = "global"))

cat(sprintf("Random transitivity:   %.4f +/- %.4f\n",
            mean(random_transitivities), sd(random_transitivities)))
cat(sprintf("Ratio Elite/Random:    %.2f\n", elite_trans / mean(random_transitivities)))
```

EliteNet's transitivity is meaningfully higher than random — triangles are  overrepresented.
Communities have structural raw material; detection is worth pursuing.

Let's use the [Louvain](https://perso.uclouvain.be/vincent.blondel/research/louvain.html) algororithm:

```{r louvain}
set.seed(123)
partition_louvain <- cluster_louvain(EliteNet)
length(partition_louvain)
```

```{r add-community-attr}
V(EliteNet)$community <- membership(partition_louvain)
head(igraph::as_data_frame(EliteNet, what = "vertices"))
```

```{r community-plot}
ggraph(EliteNet) +
  geom_edge_link(color = "grey60", alpha = 0.2) +
  geom_node_point(aes(color = factor(community)), size = 6) +
  scale_color_brewer(palette = "Dark2") +
  labs(color = "Community",
       title = sprintf("Louvain communities (n = %d)", length(partition_louvain))) +
  theme_void()
```

**Modularity Q:** if we get positive values (1 being the top value), we can consider there to be good
community structure ([wiki](https://en.wikipedia.org/wiki/Modularity_(networks))). The higher the
modularity, the denser the connections within a partition and the sparser the connections between
partitions.

```{r modularity}
modularity(EliteNet, membership(partition_louvain))
```

| Modularity Q | Interpretation                                 |
| ------------ | ------------------------------------------------ |
| < 0.3        | Weak/no meaningful communities                    |
| 0.3 - 0.7    | Moderate community structure                      |
| > 0.7        | Strong communities (possibly over-partitioned)    |



Let's prepare an ad-hoc plot. The density within a community, versus a density of the edges between them.

```{r community-density-matrix}
mem <- membership(partition_louvain)
k <- length(partition_louvain)
sz <- as.integer(table(mem))
comm_edges <- matrix(0, k, k)
el_num <- as_edgelist(EliteNet, names = FALSE)
mem_num <- as.integer(mem)
for (i in seq_len(nrow(el_num))) {
  cu <- mem_num[el_num[i,1]]; cv <- mem_num[el_num[i,2]]
  comm_edges[cu,cv] <- comm_edges[cu,cv] + 1
  comm_edges[cv,cu] <- comm_edges[cv,cu] + 1
}
dens <- matrix(0, k, k)
for (i in 1:k) {
  for (j in 1:k) {
    if (i == j) {
      dens[i,j] <- if (sz[i] > 1) comm_edges[i,j] / (sz[i] * (sz[i] - 1)) else 0
    } else {
      dens[i,j] <- comm_edges[i,j] / (sz[i] * sz[j])
    }
  }
}
dens_rounded <- round(dens, 3)
dimnames(dens_rounded) <- list(as.character(1:k), as.character(1:k))
dens_df <- as.data.frame(as.table(dens_rounded))
colnames(dens_df) <- c("comm_i", "comm_j", "density")

ggplot(dens_df, aes(x = comm_j, y = comm_i, fill = density)) +
  geom_tile(color = "white") +
  geom_text(aes(label = density), color = "white", size = 3) +
  scale_fill_viridis_c() +
  labs(x = NULL, y = NULL) +
  theme_minimal()
```

The diagonal is denser than the off-diagonal — communities are not pure fiction. But the off-diagonal is
far from empty. Let's add the `ratio`: community-internal density divided by the max density to any other
community.

```{r community-ratios}
ratios <- sapply(1:k, function(i) dens[i,i] / max(dens[i, -i]))

ratio_table <- data.frame(
  community    = 1:k,
  size         = sz,
  internal     = round(diag(dens), 3),
  max_external = round(sapply(1:k, function(i) max(dens[i, -i])), 3),
  ratio        = round(ratios, 2)
)
ratio_table
```

As expected, diagonal values are higher than off diagonals. Now you show that ratio are just above 2.3.

Notice, visualization can not be honest, you can force visuals:


```{r hive-plot, fig.width=8, fig.height=8}
V(EliteNet)$louvain <- factor(V(EliteNet)$community)
V(EliteNet)$degree  <- degree(EliteNet)

ggraph(EliteNet, layout = "hive",
       axis = louvain,
       sort.by = degree) +
  geom_edge_hive(color = "grey80", alpha = 0.25) +
  geom_axis_hive(aes(colour = louvain), label = FALSE) +
  geom_node_point(aes(colour = louvain)) +
  geom_node_text(aes(label = name), repel = TRUE, max.overlaps = 50, size = 2.5) +
  coord_fixed() +
  theme_void()
```
